\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-16 trunk/europa.tex
merged minor mods to europa secs
citations added to europa text

The \textbf {Language Modules}: Describing problem domains and problem instances in terms of objects, variables and domain rules in a programmatic way is cumbersome to the extent of being impractical, that motivated the creation of modeling languages that support those definitions in a much more natural and compact way. NDDL is the main modeling language for \eu and will be described in detail in the Modeling section below. ANML is a language under development that provides a more intuitive syntax \comment{provide ANML citation}.

@@

In an effort to mimimize the amount of work that a user needs to perform to solve specific problems, \eu  also provides extension modules that have been found to be useful in many domains: 
\begin{enumerate}
	\item \textit{Temporal Network}: Extensions to the CRE module to reason about temporal constraints 
	\item \textit{Resources}: Extensions to the CRE, PDB and Solver modules to provide out-of-the-box reasoning and search for metric resources.
	\item \textit{Chronological Backtracking Solver}: A general purpose built-in solver to provide the ability to plan out-of-the-box, problems that require challenging search may require specialized solvers to be built as explained in the Search section below.	
\end{enumerate}

@@

To understand how the most important elements of NDDL work, let's take a
look at how the Shopping Agent problem could be modeled:

@@

\begin{enumerate}
	\item \textbf{Object Types}: NDDL supports an object-oriented approach, where common data and behavior can be abstracted as a class. In this case, we have the \textit{Product} class represent the products that the Shopping Agent needs, \textit{Location} represents the places where products can be found, \textit{ProductLocation} represents the association arising when a product can be purchased at a location. The Shopping Agent itself is represented by an \textit{Agent} class, which relies on the \textit{AgentLocation} class to keep track of where the agent is at any point in time. NDDL supports a single rooted class hierarchy so that state and behavior can be extended through inheritance. We took advantage of that by making \textit{AgentLocation} inherit from \eu's built-in \textit{Timeline} class which doesn't allow predicates to overlap in time for the same object instance, in this way we can ensure that and agent can only be at one place at a time. 
	\item \textbf{Variable Types and Instances} : All object attributes are represented as variables in this model, for instance, it uses integer variables to represent the built-in temporal parameters for each token (activity duration, start and end times), string variables to represent  \textit{Location} and  \textit{Product} names, and object reference variables to represent the structural relationships between different objects (for instance, each Agent has variable to keep track of its  \textit{AgentLocation}).
	\item \textbf{Predicate Types}: Some classes like \textit{Product}, \textit{Location} and \textit{ProductLocation} are completely static and so don't have any predicate or action types associated with them. The \textit{AgentLocation} class is purely used to keep track of state and so it has \texttt{At} and \texttt{Going} predicate types associated with it. The \textit{Agent} class keeps track of the products it purchases through the \texttt{Own} predicate, having a separate \textit{AgentBag} class with its own predicates would have been an equally valid modeling choice.
	\item \textbf{Action Types}: The \textit{Agent} class is the only one that has behavior associated with it, consequently, it defines action types to represent when the agent is taking action to \texttt{Go} to a \textit{Location}, or to \texttt{Buy} a \textit{Product}. We'll see how the actual behavior for those actions is specified below.
\end{enumerate}

@@

\begin{enumerate}
  \item \textbf{Variables}: A Token Type (predicate or action) is part of a class, therefore class attributes (see references to \texttt{object.location} above), Token Type parameters (see references to \texttt{[from, to]}  in \texttt{Agent::Go}, or to \texttt{[duration, product]} in \texttt{Agent::Buy}  and locally declared variables (see \texttt{possibleStores} in \texttt{Agent::Buy}) can all be used in the definition of a Rule. 
  \item \textbf{Constraints}: restrictions on the values that variables can take inside a Rule are expressed through constraints, in the example above only the \texttt{eq} constraint is used, but in general, any constraint available in \eu's library can be used.
 \item \textbf{Subgoals}: In \texttt{Agent::Go}, subgoals are used to specify that the Shopping Agent's location must transition from origin to destination in a temporally consistent way. The \texttt{[meets, equals, metby]} operators are used to specify subgoals (Allen operators are explained in more detail below), along with the temporal relationship each token representing a subgoal has with the the token representing the \texttt{Agent::Go} action. Also, \texttt{[condition, effect]} annotations can be used when defining a subgoal inside an action, \eu doesn't immediately associate any semantics with them, but they can be extremely useful for search modules when creating plans. 
\end{enumerate}

@@

A problem instance consists mainly of 
\begin{enumerate}
  \item \textbf{Variables} and \textbf{Objects}: In this case we defined temporal variables that are used to specify the planning horizon (see \texttt{horizonStart} and \texttt{horizonEnd}), and objects are used to specify the locations, products, product-loction associations and the Shopping Agent that we want to plan for.
  \item \textbf{Facts}: These are used to specify the current state of the world and don't need any support. Facts are represented as tokens in \eu's Plan Database. In this case we use facts to specify that the Shipping Agent is initially at home.
  \item \textbf{Goals}:  Goals are represented as tokens in \eu's Plan Database, differently from Facts, Goals trigger inference in \eu which determines whether flaws (unresolved subgoals, overloaded resources, etc., flaws will be covered in detail in the search section below) need to be resolved, if that is the case, search modules can be invoked which will look for a plan where all the flaws are resolved.
  \item \textbf{Constraints}: restrictions on and between actions and goals are specified through constraints, in this case we use temporal constraints to specify when the facts hold and when the goals should be accomplished. As always, all the constraints in \eu's library (built-in, or user defined), can be used to specify these restrictions.
\end{enumerate}

@@

\comment{ add figure}

@@

\begin{enumerate}
  \item \textbf{Reservoir} resources can have both 'consume' and 'produce' tokens. Note
that these tokens do NOT have the usual start/end/duration token
variables, since occur at a single instant in time. Instead, they have a
single variable, 'time' used to represent that instant.

@@

All this information is kept in \eus Plan Database so that inference and
search mechanisms can be used to look for a problem solution.

@@

\begin{enumerate}
	\item Line \ref{li:propagate}: Propagate the constraints to test for inconsistency. If found to be inconsistent, then we can return false since this is a dead-end i.e. no refinements to p can yield a consistent plan.
	\item Line \ref{li:selectflaw}: Choose a flaw from the set of available flaws.
	\item Line \ref{li:noflaw}: If there are no flaws, then p is complete and we can terminate with success.
	\item Line \ref{li:decisionpt}: Otherwise we formulate a decision point which is a branch in the search space. Each choice is a particular refinement operation and the DecisionPoint collects all possible refinement operations for the given flaw.
	\item Line \ref{li:decisionloop}: We keep trying until the chosen flaw is resolved, or until we run out of resolvers to try.
	\item Line \ref{li:decide}: A new partial plan is obtained by application of a refinement operator. Note that the ordering over refinement operators to select is a non-deterministic step.
	\item Line \ref{li:recurse}: Recursive call to solve the new planning problem. If successful, then we are done.
	\item Line \ref{li:backtrack}: Otherwise, retract the last refinement operation and move on to try the next one.
	\item Line \ref{li:noplan}:  If we arrive here, then we have exhausted all options to resolve the flaw, including the case where no options were available initially. Thus the problem cannot be solved.
\end{enumerate}

@@

RCPSP solver:
\begin{verbatim}
    public void solve(PSEngine psengine,
    		          long timeout, // in msecs
    		          int bound)
    {
       init(psengine,timeout,bound);

>>>>

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
that motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. NDDL is the main
modeling language for \eu and will be described in detail in the
Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

In an effort to mimimize the amount of work that a user needs to
perform to solve specific problems, \eu also provides extension
modules that have been found to be useful in many domains:

\item \textit{Temporal Network}: Extensions to the CRE module to
  reason about temporal constraints

\item \textit{Resources}: Extensions to the CRE, PDB and Solver
  modules to provide out-of-the-box reasoning and search for metric
  resources.

\item \textit{Chronological Backtracking Solver}: A general purpose
  built-in solver to provide the ability to plan out-of-the-box,
  problems that require challenging search may require specialized
  solvers to be built as explained in the Search section below.
\end{enumerate}

@@

To understand how the most important elements of NDDL work, we show
how the Shopping Agent problem could be modeled:

@@

\item \textbf{Object Types}: NDDL supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. NDDL supports a single rooted class hierarchy so
  that state and behavior can be extended through inheritance. We took
  advantage of that by making \textit{AgentLocation} inherit from
  \eu's built-in \textit{Timeline} class which doesn't allow
  predicates to overlap in time for the same object instance, in this
  way we can ensure that and agent can only be at one place at a time.

\item \textbf{Variable Types and Instances} : All object attributes
  are represented as variables in this model, for instance, it uses
  integer variables to represent the built-in temporal parameters for
  each token (activity duration, start and end times), string
  variables to represent \textit{Location} and \textit{Product} names,
  and object reference variables to represent the structural
  relationships between different objects (for instance, each Agent
  has variable to keep track of its \textit{AgentLocation}).

\item \textbf{Predicate Types}: Some classes like \textit{Product},
  \textit{Location} and \textit{ProductLocation} are completely static
  and so don't have any predicate or action types associated with
  them. The \textit{AgentLocation} class is purely used to keep track
  of state and so it has \texttt{At} and \texttt{Going} predicate
  types associated with it. The \textit{Agent} class keeps track of
  the products it purchases through the \texttt{Own} predicate, having
  a separate \textit{AgentBag} class with its own predicates would
  have been an equally valid modeling choice.

\item \textbf{Action Types}: The \textit{Agent} class is the only one
  that has behavior associated with it, consequently, it defines
  action types to represent when the agent is taking action to
  \texttt{Go} to a \textit{Location}, or to \texttt{Buy} a
  \textit{Product}. We'll see how the actual behavior for those
  actions is specified below.
\end{enumerate}

@@

\begin{enumerate}
\item \textbf{Variables}: A Token Type (predicate or action) is part
  of a class, therefore class attributes (see references to
  \texttt{object.location} above), Token Type parameters (see
  references to \texttt{[from, to]} in \texttt{Agent::Go}, or to
  \texttt{[duration, product]} in \texttt{Agent::Buy} and locally
  declared variables (see \texttt{possibleStores} in
  \texttt{Agent::Buy}) can all be used in the definition of a Rule.

\item \textbf{Constraints}: restrictions on the values that variables
  can take inside a Rule are expressed through constraints, in the
  example above only the \texttt{eq} constraint is used, but in
  general, any constraint available in \eu's library can be used.

  \item \textbf{Subgoals}: In \texttt{Agent::Go}, subgoals are used to
    specify that the Shopping Agent's location must transition from
    origin to destination in a temporally consistent way. The
    \texttt{[meets, equals, metby]} operators are used to specify
    subgoals (Allen operators are explained in more detail below),
    along with the temporal relationship each token representing a
    subgoal has with the the token representing the \texttt{Agent::Go}
    action. Also, \texttt{[condition, effect]} annotations can be used
    when defining a subgoal inside an action, \eu doesn't immediately
    associate any semantics with them, but they can be extremely
    useful for search modules when creating plans.
\end{enumerate}

@@

A problem instance consists mainly of

\item \textbf{Variables} and \textbf{Objects}: In this case we defined
  temporal variables that are used to specify the planning horizon
  (see \texttt{horizonStart} and \texttt{horizonEnd}), and objects are
  used to specify the locations, products, product-location
  associations and the Shopping Agent that we want to plan for.

\item \textbf{Facts}: These are used to specify the current state of
  the world and don't need any support. Facts are represented as
  tokens in \eu's Plan Database. In this case we use facts to specify
  that the Shipping Agent is initially at home.

\item \textbf{Goals}: Goals are represented as tokens in \eu's Plan
  Database, differently from Facts, Goals trigger inference in \eu
  which determines whether flaws (unresolved subgoals, overloaded
  resources, etc., flaws will be covered in detail in the search
  section below) need to be resolved, if that is the case, search
  modules can be invoked which will look for a plan where all the
  flaws are resolved.

\item \textbf{Constraints}: restrictions on and between actions and
  goals are specified through constraints, in this case we use
  temporal constraints to specify when the facts hold and when the
  goals should be accomplished. As always, all the constraints in
  \eu's library (built-in, or user defined), can be used to specify
  these restrictions.
\end{enumerate}

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (STN)]{\label{fig:stn1}\includegraphics[width=0.4\columnwidth]{figs/stn-graph.pdf}}
  \subfloat[\small Resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.4\columnwidth]{figs/distance-graph.pdf}}
    \caption{\small An illustration of a STN and its corresponding
      distance graph.}
  \label{fig:stn}
\end{figure}

% \comment{ add figure. Use Europa wiki figure; or use conor figure from slides}

@@

\item \textbf{Reservoir} resources can have both 'consume' and
  'produce' tokens. Note that these tokens do NOT have the usual
  start/end/duration token variables, since occur at a single instant
  in time. Instead, they have a single variable, 'time' used to
  represent that instant.

@@

This information is kept in \eus Plan Database so that inference and
search mechanisms can be used to look for a problem solution.

@@

\item Line \ref{li:propagate}: Propagate the constraints to test for
  inconsistency. If found to be inconsistent, then we can return false
  since this is a dead-end i.e. no refinements to p can yield a
  consistent plan.

\item Line \ref{li:selectflaw}: Choose a flaw from the set of
  available flaws.

\item Line \ref{li:noflaw}: If there are no flaws, then p is complete
  and we can terminate with success.

\item Line \ref{li:decisionpt}: Otherwise we formulate a decision
  point which is a branch in the search space. Each choice is a
  particular refinement operation and the DecisionPoint collects all
  possible refinement operations for the given flaw.

\item Line \ref{li:decisionloop}: We keep trying until the chosen flaw
  is resolved, or until we run out of resolvers to try.

\item Line \ref{li:decide}: A new partial plan is obtained by
  application of a refinement operator. Note that the ordering over
  refinement operators to select is a non-deterministic step.

\item Line \ref{li:recurse}: Recursive call to solve the new planning
  problem. If successful, then we are done.

\item Line \ref{li:backtrack}: Otherwise, retract the last refinement
  operation and move on to try the next one.

\item Line \ref{li:noplan}: If we arrive here, then we have exhausted
  all options to resolve the flaw, including the case where no options
  were available initially. Thus the problem cannot be solved.
\end{enumerate}

@@

RCPSP solver \comment{Do we need this example?}:
\begin{verbatim}
    public void solve(PSEngine psengine,
    		          long timeout, // in msecs
    		          int bound)
    {
       init(psengine,timeout,bound);
